---
title: "MANOJ FML ASSIGNMENT 2"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem Statement

  Universal bank is a young bank growing rapidly in terms of overall customer acquisition. The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.  
  
  A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.
  
  The file UniversalBank.csv contains data on 5000 customers. The data include customer demographic information (age, income, etc.), the customer’s relationship with the bank (mortgage, securities account, etc.), and the customer response to the last personal loan campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the personal loan that was offered to them in the earlier campaign.
  
Split the data into training (60%) and validation (40%) sets.
  
Load the libraries which are required for the code

```{r}
library(class)
library(caret)
library(e1071)
library(knitr)
```

Read the data from UniversalBank.csv 

```{r}
universalbank1.df <- read.csv("C:\\Users\\yadla sreebhavya\\Downloads\\UniversalBank (2).csv")
dim(universalbank1.df)
t(t(names(universalbank1.df))) 
# The 't' function used to transpose the given data frame
```

```{r}
universalbank1.df <- universalbank1.df[,-c(1,5)]
```

Transform categorical variables into dummy variables

```{r}
# Education needs to be converted to factor for the data

universalbank1.df$Education <- as.factor(universalbank1.df$Education)

```

```{r}
# Transform Education into Dummy Variables

groups <- dummyVars(~., data = universalbank1.df) # This creates the dummy groups

universal_m.df <- as.data.frame(predict(groups,universalbank1.df))
```

The data will be splitted into 60% training and 40 % Validation 

```{r}
set.seed(1)  
# It is crucial to guarantee that we obtain an identical sample when we re execute the code.
train.index <- sample(row.names(universal_m.df), 0.6*dim(universal_m.df)[1])
valid.index <- setdiff(row.names(universal_m.df), train.index)  
train.df <- universal_m.df[train.index,]
valid.df <- universal_m.df[valid.index,]
t(t(names(train.df)))

```

### Now normalize the data

```{r}
train.norm.df <- train.df[,-10] 
# Income is the 10th variable
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
```

Question 1

Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

We have transformed all dummy variables from categorical variables

 
```{r}

new_customer1 <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1)
```

Normalize the new customer

```{r}
new.cust.norm <- new_customer1
new.cust.norm <- predict(norm.values,new.cust.norm)

```

Now let us predict using K-NN

```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1

```
Question 2

What is a choice of k that balances between over fitting and ignoring the predictor information?
Calculate the accuracy for k    


```{r}

accuracy.df <- data.frame(k = seq(1, 20, 1), overallaccuracy = rep(0, 20)) 
for(i in 1:20) 
  {knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred,as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
}
which(accuracy.df[,2] == max(accuracy.df[,2])) 
plot(accuracy.df$k,accuracy.df$overallaccuracy, main = "Accuracy Vs K", xlab = "k", ylab = "accuracy")

```
Question 3

Show the confusion matrix for the validation data that results from using the best k.

Perform Confusion Matrix for best K=3

```{r}

knn.pred <- class::knn(train = train.norm.df,
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = 3)

confusionMatrix(knn.pred,as.factor(valid.df$Personal.Loan))

```
Question 4

Consider the following customer: Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.

#New customer profile

```{r}
new_customer2<-data.frame(
  Age = 40, 
  Experience = 10, 
  Income = 84, 
  family =2, 
  CCAvg = 2, 
  Education_1 = 0,
  Education_2 = 1, 
  Education_3 = 0, 
  Mortgage = 0, 
  Securities.Account = 0, 
  CDAccount = 0, 
  Online = 1, 
  CreditCard = 1)

```


```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 3)
knn.pred1
```

Print the predicted class (1 for Approval , 0 for Not Approval)

```{r}

print("This customer is classified as:Not Approval")

```
Question 5

Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.
    

Seperate the data to 50% training and 30% Validation and 20% Testing 

```{r}
set.seed(1)
Train_Index1 <- sample(row.names(universal_m.df), 0.5*dim(universal_m.df)[1])
Val_Index1 <- sample(setdiff(row.names(universal_m.df),Train_Index1),0.3*dim(universal_m.df)[1])
Test_Index1 <-setdiff(row.names(universal_m.df),union(Train_Index1,Val_Index1))
Train_Data <- universal_m.df[Train_Index1,]
Validation_Data <- universal_m.df[Val_Index1,]
Test_Data <- universal_m.df[Test_Index1,]
```

Normalize the data

```{r}
train.norm.df1 <- Train_Data[,-10]
valid.norm.df1 <- Validation_Data[,-10]
Test.norm.df1  <-Test_Data[,-10]

norm.values1 <- preProcess(Train_Data[, -10], method=c("center", "scale"))
train.norm.df1 <- predict(norm.values1, Train_Data[,-10])
valid.norm.df1 <- predict(norm.values1, Validation_Data[,-10])
Test.norm.df1 <-predict(norm.values1,Test_Data[,-10])

```

Predict using K-NN

```{r}

validation_check= class::knn(train = train.norm.df1, 
                           test = valid.norm.df1,  
                           cl = Train_Data$Personal.Loan, 
                           k = 3)

test_check = class::knn(train = train.norm.df1, 
                     test = Test.norm.df1,  
                     cl = Train_Data$Personal.Loan, 
                     k = 3)

Train_check = class::knn(train = train.norm.df1, 
                     test = train.norm.df1,  
                     cl = Train_Data$Personal.Loan, 
                     k = 3)
```

Validation confusion Matrix

```{r}
validation_conf_matrix = confusionMatrix(validation_check, 
                                               as.factor(Validation_Data$Personal.Loan), 
                                               positive = "1")

validation_conf_matrix

```
### Test confusion Matrix

```{r}

test_conf_matrix = confusionMatrix(test_check, 
                                         as.factor(Test_Data$Personal.Loan), 
                                         positive = "1")


test_conf_matrix

```

```{r}

Training_conf_matrix = confusionMatrix(Train_check, 
                                               as.factor(Train_Data$Personal.Loan), 
                                               positive = "1")

Training_conf_matrix
```


Discrepancies in Model Evaluation Metrics

Training vs. Testing:

Accuracy: The training dataset displays a higher accuracy score (0.9764) compared to the test dataset (0.961).

Explanation: This variance in accuracy can be attributed to disparities in the datasets used for evaluation. The training data may be more balanced or contain patterns that are easier for the model to grasp.

Sensitivity (True Positive Rate): The training dataset demonstrates greater sensitivity (0.7672) than the test dataset (0.6875).

Explanation: This indicates that the model trained on the training data excels at correctly identifying positive cases, such as loan approvals. It likely has a reduced rate of false negatives.

Specificity (True Negative Rate): The training dataset showcases higher specificity (0.9978) compared to the test dataset (0.9955).

Explanation: This implies that the model trained on the training data is better at accurately recognizing negative cases, like loan rejections. It may have a lower rate of false positives.

Positive Predictive Value (Precision): The training dataset boasts a higher positive predictive value (0.9727) compared to the test dataset (0.9506).

Explanation: The model trained on the training data excels in precision when predicting positive cases, resulting in fewer false positive predictions.

Training vs. Validation:

Accuracy: The training dataset still maintains a superior accuracy score (0.968) in contrast to the validation dataset (0.961).

Explanation: Similar to the comparison with the test dataset, the training data may possess characteristics that make it more conducive for the model to achieve a high level of accuracy.

Sensitivity (True Positive Rate): The training dataset exhibits higher sensitivity (0.69118) relative to the validation data set (0.6875).

Explanation: The model trained on the training data is better at accurately identifying positive cases. This suggests that the validation data set model might have a higher rate of false negatives.

Specificity (True Negative Rate): The training dataset still presents higher specificity (0.99560) compared to the validation data set (0.9955).

Explanation: The model trained on the training data excels at accurately recognizing negative cases. The validation data set model might have a slightly higher rate of false positives.

Positive Predictive Value (Precision): The training dataset continues to possess a greater positive predictive value (0.9506) compared to the validation dataset (0.9400).

Explanation: The model trained on the training data maintains its precision in predicting positive cases, resulting in fewer false positive predictions.

#Potential Reasons for Discrepancies:

Data Set Differences: Variations in data composition and distribution among different sets can significantly influence model performance. For instance, one data set may be more imbalanced, making it challenging to predict rare events.

Model Variability: Differences in model configurations or the random initialization of model parameters can lead to variations in performance.
 .
 